STAT 450 Project: Real Estate

### 2020-03-20 src/2nd_draft.Rmd
eda updated

---
title: "2nd_draft"
author: "Xuechun Lu, Yuting Wen, Peter Han, Yuetong Liu"
date: "15/03/2020"
output:
  github_document
---

@@

Oo reduce the dimension of our data, we aggregate all properties in the same region, classcode, and year are aggregate into a group, becasue these properties have same mill rate,  which is our response variable. Here is the summary statistics for these groups:

@@

## TODO modify this part based on the comment in the following link (Leonie)
## _______________________________
https://github.com/STAT450-550/RealEstate/blob/450/doc/Peter_Han_Individual_edit.pdf
### Measure of goodness of fit and prediction power

@@

### Cross Validation
To examine the goodness of fit and prediciton power, a 50-run of 10-fold cross validation would be performed in this study. For each run, a 10-fold corss validation was used to train the five models and make predictions on training and testing data respectively. Then, the MSEs calculated from the training data and MSPEs calculated from the testing data were stored in vectors of corresponding mdoels.

@@

##_____________________________________________

@@

@@

@@

# log transformation and take avg
assessment_transform[,9] <- assessment_transform$assessTotal/assessment_transform$propertyCount
names(assessment_transform)[9] <- paste("AvgTotalAssessmentValue")

cor(assessment_transform$rate,assessment_transform$AvgTotalAssessmentValue)

assessment_transform %>%
                   ggplot(aes(x=assessTotal,
                              y=rate,group=AddressAssessorMunicipalityDesc,
                              color=AddressAssessorMunicipalityDesc)) + 
                   geom_point() + 
                   ggtitle(sprintf("mill rate vs total assessment value")) + 
                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")

assessment_transform %>%
                   ggplot(aes(x=landTotal,
                              y=rate,group=AddressAssessorMunicipalityDesc,
                              color=AddressAssessorMunicipalityDesc)) + 
                   geom_point() + 
                   ggtitle(sprintf("mill rate vs total land assessment value")) + 
                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
assessment_transform %>%
                   ggplot(aes(x=improvementTotal,
                              y=rate,group=AddressAssessorMunicipalityDesc,
                              color=AddressAssessorMunicipalityDesc)) + 
                   geom_point() + 
                   ggtitle(sprintf("mill rate vs total improvement assessment value")) + 
                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")

assessment_transform %>%
                   ggplot(aes(x=AvgTotalAssessmentValue,
                              y=rate,group=AddressAssessorMunicipalityDesc,
                              color=AddressAssessorMunicipalityDesc)) + 
                   geom_point() + 
                   ggtitle(sprintf("mill rate vs average total assessment value")) + 
                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")

# ______________________________________________________________________________________________
# CATEGORICAL variables
for(i in colnames(assessment_transform)[2:3]){
  print(ggplot(assessment_aggregate, aes(x=as.factor(!!as.name(i)),y=rate, fill=as.factor(!!as.name(i)))) + 
  geom_boxplot()+ 
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none"))
}

# kruskal.test
kruskal.test(assessment_transform$rate~assessment_transform$TaxClassCode,data=assessment_transform)
kruskal.test(assessment_transform$rate~assessment_transform$AddressAssessorMunicipalityDesc,data=assessment_transform)
```

## ___________________________
## TODO KW result table, explanation of all plots (Yuting)
## ___________________________

##_____________________________
## TODO modify this part using boxplot and tables, change desrciptions (Leonie)
###Linear Model

**Table 2: Result from Model Used**

<!-- Model | Mutiple R-Squared | Adjusted R_Squared | MSE | PMSE  -->
<!-- ------|--------|--------|--------|-------- -->
<!--       OLR full | 0.8874| 0.8707| 1.9843| 2.5902 -->
<!--       OLR reduced| 0.8874| 0.8721| 1.9845| 2.5237 -->
<!--       Ridge|  0.8868| NA| 1.9896| 2.5675 -->
<!--       LASSO| 0.8873| NA| 1.9855| 2.5280 -->
<!--       Elastic Net| 0.8625| NA| 2.2366| 2.5486 -->

@@

For a dtailed look of how the models were fitted and evaluated, please refer to the appendix.

@@

boxplot(mspe.la, 
        mspe.ri,
        mspe.reduced, 
        mspe.f, 
       mspe.n,
        names = c("LASSO", 
                  "Ridge", 
                  "Transformed", 
                  "Full",
                  "Null"
                  ), 
        col = c("steelblue","gray80", 
                "pink",
                "tomato", "springgreen"), 
        cex.axis = 1, 
        cex.lab = 1, 
        cex.main = 2,
        main = "Prediction Power",
        ylab = expression(hat(MSPE)))
```
##_______________________________________

##_______________________________________
##TODO: make prediction work for 2020 data (Yuetong)
##(it's missing AvgTotalAssessmentValue, looks like we need to calculate them)
```{r fit, echo=FALSE}
## choose transformed model
linear_1<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(TaxClassCode)+AvgTotalAssessmentValue,data=assessment_transform)

@@

##——————————————————————————
##TODO: more styling (Yuting)
##___________________________
## Appendix

##__________________________
## TODO: add one more section: (Yuetong) 
## Include the code relevant to summarizing and aggregating the data from the client, which produces in the dataframe used for analysis.
##__________________________

>>>>

---
title: 'STAT 450 Project: Real Estate'
author: "Xuechun Lu, Yuting Wen, Peter Han, Yuetong Liu"
date: "15/03/2020"
output:
  pdf_document
---

@@

To reduce the dimension of our data, we aggregate all properties in the same region, tax class code, and year are aggregate into a group, becasue these properties have same mill rate,  which is our response variable. Here is the summary statistics for these groups:

@@

### Measure of goodness of fit and prediction power

@@

### Cross Validation

To examine the goodness of fit and prediciton power, a 50-run of 10-fold cross validation would be performed in this study. For each run, a 10-fold corss validation was used to train the five models and make predictions on training and testing data respectively. Then, the MSEs calculated from the training data and MSPEs calculated from the testing data were stored in vectors of corresponding mdoels.

@@

@@

#### Continuous Variables

@@

Since mill rate is mainly affected by total assessment, which is sum of total land assessment and total improvement assessment, scatter plots between mill rate and all kinds of assessment among different municipalities are created to show the relationship in Figure 1, Figure 2, Figure 3

@@

# transformation: taking avg
assessment_transform[,9] <- assessment_transform$assessTotal/assessment_transform$propertyCount
names(assessment_transform)[9] <- paste("AvgTotalAssessmentValue")

# rename column for better visulization
assessment_eda <- assessment_transform
names(assessment_eda)[2] <- paste("Municipality")
names(assessment_eda)[3] <- paste("Tax Class")
names(assessment_eda)[4] <- paste("Total Assessment")
names(assessment_eda)[5] <- paste("Total Land Assessment")
names(assessment_eda)[6] <- paste("Total Improvement Assessment")
names(assessment_eda)[8] <- paste("Mill Rate")
names(assessment_eda)[9] <- paste("Average Total Assessment")

assessment_eda %>%
   ggplot(aes(x=`Total Assessment`,
              y=`Mill Rate`,group=Municipality,
              color=Municipality)) + 
   geom_point() + 
   ggtitle(sprintf("Figure 1: Mill Rate vs Total Assessment")) + 
   geom_smooth(aes(group = 1), size = 0.5, method = "glm", se = FALSE, colour = "black") +
   theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

assessment_eda %>%
   ggplot(aes(x=`Total Land Assessment`,
              y=`Mill Rate`,group=Municipality,
              color=Municipality)) + 
   geom_point() + 
   ggtitle(sprintf("Figure 2: Mill Rate vs Total Land Assessment")) + 
   geom_smooth(aes(group = 1), size = 0.5, method = "glm", se = FALSE, colour = "black") +
   theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

assessment_eda %>%
   ggplot(aes(x=`Total Improvement Assessment`,
              y=`Mill Rate`,group=Municipality,
              color=Municipality)) + 
   geom_point() + 
   ggtitle(sprintf("Figure 3: Mill Rate vs Total Improvement Assessment")) + 
   geom_smooth(aes(group = 1), size = 0.5, method = "glm", se = FALSE, colour = "black") +
   theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
```

There is no clear trend between mill rate and any kind of assessment. Plots also have shown that most points are condensed on the left horizontal axis, since some municipalities have large assessment.

To reduce the effect of large assessment of some municipalities, total assessment across municipalities are transformed by taking total assessment dividing by number of properties of each municipality and tax class. The transformed data is named **"Average Total Assessment"**. A scatter plot between mill rate and average total assessment is showed in Figure 4.

```{r, echo=FALSE}
assessment_eda %>%
 ggplot(aes(x=`Average Total Assessment`,
            y=`Mill Rate`,group=Municipality,
            color=Municipality)) + 
 geom_point() + 
 ggtitle(sprintf("Figure 4: Mill rate v.s. Average Total Assessment")) + 
 geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") +
theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
```

Plot from Figure 4 has shown that mill rate tends to decrease as average total assessment increase. Also they have moderately strong linear correlation.

#### Categorical Variables

Here categorical variables were taken into account, boxplots of mill rate across municipalities and tax classes are plotted to display the distributions in Figure 5 and Figure 6.

```{r, echo=FALSE}
# ______________________________________________________________________________________________
# CATEGORICAL variables
assessment_eda %>%
     ggplot(aes(x=Municipality,
                y=`Mill Rate`,fill=Municipality)) + 
     geom_boxplot() + 
     ggtitle(sprintf("Figure 5: Mill rate across Municipalities")) +
     theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")

assessment_eda %>%
     ggplot(aes(x=as.factor(`Tax Class`),
                y=`Mill Rate`,fill=`Tax Class`)) + 
     geom_boxplot() + 
     ggtitle(sprintf("Figure 6: Mill rate across Tax Classes")) +
     theme(legend.position = "none") +
     labs(x="Tax Class") +
     scale_x_discrete(labels=c("1", "5", "6"))
```

Figure 5 has shown that most municipalities have different mean and variance in terms of mill rates.

Figure 6 also supports that there is unequal mean and variance across tax classes.

A statistical test called Kruskal-Wallis Test is also performed to test their distribution. It is used to decide if population distributions are identical, and the corresponding p-value which is smaller than 0.05 can indicate that the data has nonidentical distributions.

The results of Kruskal-Wallis Test is shown in Table 1.

**Table 1: Kruskal-Wallis Test of Mill Rate across Municipality and Tax Class**

Distribution  | p-value
------|--------
Mill Rate across Municipality | 0.00675
Mill Rate across Tax Class | < 2.2e-16

The p-values in Table 1 supports that there is nonidentical distributions of mill rates across municipalities and tax classes.

```{r, echo=FALSE, results='hide'}
# kruskal.test
kruskal.test(assessment_transform$rate~assessment_transform$TaxClassCode,data=assessment_transform)
kruskal.test(assessment_transform$rate~assessment_transform$AddressAssessorMunicipalityDesc,data=assessment_transform)
```

\pagebreak
###Linear Model

@@

For a dtailed look of how the models were fitted and evaluated, please refer to the **Appendix**.

@@

boxplot(mspe.la, 
        mspe.ri,
        mspe.reduced, 
        mspe.f, 
       mspe.n,
        names = c("LASSO", 
                  "Ridge", 
                  "Transformed", 
                  "Full",
                  "Null"
                  ), 
        col = c("steelblue","gray80", 
                "pink",
                "tomato", "springgreen"), 
        cex.axis = 1, 
        cex.lab = 1, 
        cex.main = 2,
        main = "Prediction Power",
        ylab = expression(hat(MSPE)))
```

```{r fit, echo=FALSE}
## choose transformed model
linear_1<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(TaxClassCode)+AvgTotalAssessmentValue,data=assessment_transform)

@@

## Appendix
